### 2022-11-16 austins_sim.rmd
Adding main and austins_sim .rmd files

#doc for Austin to create simulations

```{r}
library(tidyverse)
library(dplyr)
```

```{r}
data <- read_csv("owid-covid-data.csv")
```

```{r}
names(data)
china <- data %>% filter(location == "China", date < "2020-03-15")
```

```{r}
cases <- ggplot(china, aes(x = date, y = total_cases)) + geom_line() + labs(title = "Total Cases in China", x = "Date", y = "Total Cases")
fatalities <- ggplot(china, aes(x = date, y = total_deaths)) + geom_line() + labs(title = "Total Deaths in China", x = "Date", y = "Total Deaths")
```

```{r}
cases
fatalities
```

```{r}
```