### 2022-11-22 austins_sim.rmd
Added Error Functions to Graphs

#doc for Austin to create simulations

```{r}
library(tidyverse)
library(dplyr)
```

```{r}
data <- read_csv("owid-covid-data.csv")
```

```{r}
names(data)
china <- data %>% filter(location == "China", date < "2020-03-20")
```

```{r}
function <- (5) + (10)*erf((10)*x-14)
```

```{r}
cases <- ggplot(china, aes(x = date, y = total_cases)) + 
    geom_line() + 
    labs(title = "Total Cases in China", x = "Date", y = "Total Cases")
fatalities <- ggplot(china, aes(x = date, y = total_deaths)) + 
    geom_line() + 
    labs(title = "Total Deaths in China", x = "Date", y = "Total Deaths")
```

```{r}
cases
fatalities
```

```{r}
```