### 2022-11-30 austins_sim.rmd
Added Powerpoint rmd and continued analysis

#doc for Austin to create simulations

```{r}
library(tidyverse)
library(dplyr)
library(pracma)
```

```{r}
data <- read_csv("owid-covid-data.csv")
```

```{r}
names(data)
china <- data %>% filter(location == "China", date > "2020-01-22", date < "2020-03-23")
china <- subset(china, select = c(date, total_cases, total_deaths))
china <- china %>% mutate(days = 1:n())
```

```{r}
cases_China <- ggplot(china, aes(x = date, y = total_cases)) + 
    geom_line() + 
    labs(title = "Total Cases in China", x = "Date", y = "Total Cases")
fatalities_China <- ggplot(china, aes(x = date, y = total_deaths)) + 
    geom_line() + 
    labs(title = "Total Deaths in China", x = "Date", y = "Total Deaths")
print(cases_China)
print(fatalities_China)
```

CHINA SIMULATION
```{r}
x <- 0:60
cases_function_china <- function(x) 40000+40000*erf(0.11*x-2)
fatalities_function_china <- function(x) 1600+1600*erf(0.070*x-1.7)

cases_plot_China <- ggplot(china, aes(x = days, y = total_cases)) + 
    geom_line() + 
    stat_function(fun=cases_function_china, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 60, 1)) +
    labs(title = "Total Cases in China from 1/22/2020 to 3/23/2020 (60 days)", x = "Days", y = "Total Cases")
fatalities_plot_China <- ggplot(china, aes(x = days, y = total_deaths)) + 
    geom_line() + 
    stat_function(fun=fatalities_function_china, color="#3333e2", size=1.5) +
    scale_x_continuous(breaks = seq(0, 60, 1)) +
    labs(title = "Total Deaths in China from 1/22/2020 to 3/23/2020 (60 days)", x = "Days", y = "Total Deaths")
print(cases_plot_China)
print(fatalities_plot_China)
```

AUSTRALIA SIMULATION
```{r}
australia <- data %>% filter(location == "Australia", date > "2020-06-01", date < "2020-12-19")
australia <- subset(australia, select = c(date, total_cases, total_deaths))
australia <- australia %>% mutate(days = 1:n())
```

```{r}
minimizedDays <- 70
australiaPre <- australia %>% filter(days < minimizedDays)
cases_Australia <- ggplot(australiaPre, aes(x = date, y = total_cases)) + 
    geom_line() + 
    labs(title = "Total Cases in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Date", y = "Total Cases")
fatalities_Australia <- ggplot(australiaPre, aes(x = date, y = total_deaths)) +
    geom_line() + 
    labs(title = "Total Deaths in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Date", y = "Total Deaths")
print(cases_Australia)
print(fatalities_Australia)
```

```{r}
x <- 0:minimizedDays
cases_function_australia <- function(x) 17200+10000*erf(0.04*x-2.4)
fatalities_function_australia <- function(x) 500+400*erf(0.038*x-3.1)

cases_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_line() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=2, linetype="dotted") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 1)) +
    labs(title = "Total Cases in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Cases")
fatalities_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_line() + 
    stat_function(fun=fatalities_function_australia, color="#3333e2", size=2, linetype="dotted") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 1)) +
    labs(title = "Total Deaths in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Deaths")
print(cases_pre_plot_Australia)
print(fatalities_pre_plot_Australia)
```

```{r}
x <- 0:200
cases_function_australia <- function(x) 17200+10000*erf(0.04*x-2.4)
fatalities_function_australia <- function(x) 500+400*erf(0.038*x-3.1)

cases_plot_Australia <- ggplot(australia, aes(x = days, y = total_cases)) + 
    geom_line() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 1)) +
    labs(title = "Total Cases in Australia from 6/1/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Cases")
fatalities_plot_Australia <- ggplot(australia, aes(x = days, y = total_deaths)) +
    geom_line() + 
    stat_function(fun=fatalities_function_australia, color="#3333e2", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 1)) +
    labs(title = "Total Deaths in Australia from 6/1/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Deaths")

print(cases_plot_Australia)
print(fatalities_plot_Australia)
```

```{r}

```