### 2022-11-30 jin_sim.Rmd
Adding Jin's simulation file.

---
title: "jin_sim"
output: html_document
date: '2022-11-29'
---

## load library

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(pracma)
```

## read data

```{r}
data <- read.csv("https://raw.githubusercontent.com/stat400-csu/project-5/main/owid-covid-data.csv")
View(data)
```

## extract China data of interest

```{r}
china <- data |>
  filter(location == "China") |>
  filter(date >= "2020-01-22", date <= "2020-03-28") |>
  select(c(date, total_cases, total_deaths)) |>
  mutate(days = 1:n())
head(china)
```

## recreate cumulative number plots of positive cases and fatalities in China

```{r}
positive <- ggplot(data = china) +
  geom_point(aes(x = days, y = total_cases), color = "red") +
  labs(x = "Days", y = "China cumulative number of cases")
positive
fatalities <- ggplot(data = china) +
  geom_point(aes(x = days, y = total_deaths), color = "red") +
  labs(x = "Days", y = "China cumulative number of fatalities") +
  ylim(0,5000)
fatalities
```

## apply nonlinear least square curve fitting to find the targeted guassian error function

```{r}
gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
model1 <- nls(total_cases ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
summary(model1)
model2 <- nls(total_deaths ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
summary(model2)
```

## superimpose the fitted gaussian error functions to the plots

```{r, warning=FALSE}
fn_model1 <- function(x) {4.017e+04 + 4.047e+04 * erf(9.284e-02 * x - 1.732e+00)}
positive +
  stat_function(fun=fn_model1, color="blue")
fn_model2 <- function(x) {1.571e+03 + 1.658e+03 * erf(6.194e-02 * x - 1.533e+00)}
fatalities +
  stat_function(fun=fn_model2, color="blue")
```

## Monte Carlo simulatioin design

```{r}
set.seed(400)

n <- length(china$days)   # number of observed days
m <- 150    # number of random outcomes
mu <- 1
sig <- 0.01  # the assumed measurement uncertainty is 10%
gau_sample <- matrix(rnorm(n*m, mu, sig), nrow=m, ncol=n)   # matrix of gaussian distribution

# multiply the nominal value with gau_sample for each day
mc_data <- vector()
for (j in 1:n){
  mc_data <- c(mc_data, china$total_cases[j]*gau_sample[,j])
}
mc_data <- matrix(mc_data, nrow=m, ncol=n)  # daily data to obtain 150 series of cases
head(mc_data)
```

## Day of flex (found with the fitted model)

```{r}
# trial 1
try <- mc_data[1,]
model_try <- nls(try ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
est_try <- summary(model_try)$parameters
est_try <- est_try[, 1]
names(est_try) <- NULL
fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

# find the second derivative of fitted function
# source: https://proofwiki.org/wiki/Derivative_of_Error_Function
erf_deriv <- function(x) {2/sqrt(pi)*exp(-x^2)}
fn_try_deriv <- bquote(.(est_try[2]) * 2/sqrt(pi) * exp(-(.(est_try[3]) * x - .(est_try[4]))^2))
fn_try_deriv <- D(fn_try_deriv,'x')
x=1:67
tmp <- eval(fn_try_deriv)
min(which(tmp < 0))
```

```{r}
# trial 1
try <- mc_data[1,]
model_try <- nls(try ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
est_try <- summary(model_try)$parameters
est_try <- est_try[, 1]
names(est_try) <- NULL
fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

# find the second derivative of fitted function
# source: https://proofwiki.org/wiki/Derivative_of_Error_Function
erf_deriv <- function(x) {2/sqrt(pi)*exp(-x^2)}
fn_try_deriv <- bquote(.(est_try[2]) * 2/sqrt(pi) * exp(-(.(est_try[3]) * x - .(est_try[4]))^2))
fn_try_deriv <- D(fn_try_deriv,'x')
x=1:67
tmp <- eval(fn_try_deriv)
min(which(tmp < 0))
```

## Day of flex (found without the fitted model)

```{r}
flex <- rep(NA, m)

# for each Monte Carlo simulation (m)
for (i in 1:m){
  temp_data <- mc_data[i,]
  num_diff <- 0
  # for each day
  for (j in 1:(n-1)){
    num_diff <- temp_data[j+1] - temp_data[j]
    if (num_diff < 0){
      flex[i] <- j
      break
    }
  }
}

for (i in 1:m){
  temp_data <- mc_data[i,]
  num_diff <- rep(NA,n-1)
  # for each day
  for (j in 1:(n-1)){
    num_diff[j] <- temp_data[j+1] - temp_data[j]
  }
  for (j in 1:(n-1)){
    if (num_diff[j+1] < num_diff[j]){
      flex[i] <- j
      break
    }
  }
}
flex
```

```{r}
temp_data <- mc_data[3,]
  # for each day
for (j in 1:(n-1)){
    num_diff[j] <- temp_data[j+1] - temp_data[j]
}

for (j in 1:(n-1)){
  if (num_diff[j+1] < num_diff[j]){
    print(j)
    break
  }
}
```