### 2022-12-06 paper.rmd

---
title: "Paper"
author: "Austin Lackey, Chien Lin and Jin Peng"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Introduction
### Out motivtion for this project is to...

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# useful libraries
library(tidyverse)
library(pracma)
library(minpack.lm)
# read covid data
data <- read_csv("owid-covid-data.csv")
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

```{r}
print(head(data))
```

# Insert More Below....